---
title: "geminiEDA-usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geminiEDA-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(dplyr)
library(gtsummary)
library(geminiEDA)
library(ggthemes) 
library(scales)   # Add this line
```

# Introduction

The `geminiEDA` package provides tools for generating exploratory data analysis (EDA) reports in R. This vignette demonstrates how to use the package to apply custom themes, create plots, and generate summary tables.

## Custom GEMINI Theme

The `geminiEDA` package includes a custom theme for `ggplot2` plots. You can apply this theme to your plots using the `theme_GEMINI` function.

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(geminiEDA)

# Create a sample plot with the default theme
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
p
```

Apply the GEMINI theme

```{r, message=FALSE, warning=FALSE}
p + theme_GEMINI()
```

## Custom Color and Fill Scales

The package also provides custom color and fill scales for `ggplot2` plots.

```{r, message=FALSE, warning=FALSE}
# Create a sample plot with default scales
p <- ggplot(mtcars, aes(x = factor(cyl), fill = factor(cyl))) +
  geom_bar()
p
```

Apply the GEMINI fill scale
```{r,  message=FALSE, warning=FALSE}
p + scale_fill_GEMINI()
```

Apply the GEMINI theme + fill scale
```{r,  message=FALSE, warning=FALSE}
p + 
  scale_fill_GEMINI() +
  theme_GEMINI()
```
## Univariate Plot Function

The `eda_univ_autoplot` function creates univariate plots based on the variable type.

### Numeric variable
```{r, message=FALSE, warning=FALSE}
eda_univ_autoplot(mtcars, "mpg")
```
### Categoric variable
```{r, message=FALSE, warning=FALSE}
mtcars$cyl <- factor(mtcars$cyl)
eda_univ_autoplot(mtcars, "cyl")
```
## Bivariate Plot Function

The `eda_biv_autoplot` function creates bivariate plots based on the types of the covariate and outcome variables.
### Numeric x Numeric
```{r, message=FALSE, warning=FALSE}
eda_biv_autoplot(mtcars, "wt", "mpg")
```
### Categorical vs. Categorical
```{r, message=FALSE, warning=FALSE}
eda_biv_autoplot(mtcars, "cyl", "am")
```
### Categorical vs. Numeric
```{r, message=FALSE, warning=FALSE}
eda_biv_autoplot(mtcars, "cyl", "mpg")
```
## Summary Table Functions

The package includes functions to generate summary tables for univariate and bivariate analyses.
### Univariate tabke
```{r, message=FALSE, warning=FALSE}
eda_univ_summary(mtcars, "mpg")
```

### Bivariate table
```{r, message=FALSE, warning=FALSE}
eda_biv_summary(mtcars, "cyl", "mpg")
```
# Conclusion

This vignette provided an overview of the `geminiEDA` package and demonstrated how to use its functions to create plots and generate summary tables for EDA. For more information, please refer to the package documentation.
