---
title: "geminiEDA-usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{geminiEDA-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(dplyr)
library(gtsummary)
library(geminiEDA)
library(ggthemes) 
library(scales)   # Add this line
```

# Introduction

The `geminiEDA` package provides tools for generating exploratory data analysis (EDA) reports in R. This vignette demonstrates how to use the package to apply custom themes, create plots, and generate summary tables.

## Custom GEMINI Theme

The `geminiEDA` package includes a custom theme for `ggplot2` plots. You can apply this theme to your plots using the `theme_GEMINI` function.

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(geminiEDA)

# Create a sample plot with the default theme
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
p
```

Apply the GEMINI theme

```{r, message=FALSE, warning=FALSE}
p + theme_GEMINI()
```

## Custom Color and Fill Scales

The package also provides custom color and fill scales for `ggplot2` plots.

```{r, message=FALSE, warning=FALSE}
# Create a sample plot with default scales
p <- ggplot(mtcars, aes(x = factor(cyl), fill = factor(cyl))) +
  geom_bar()
p
```

Apply the GEMINI fill scale
```{r,  message=FALSE, warning=FALSE}
p + scale_fill_GEMINI()
```

Apply the GEMINI theme + fill scale
```{r,  message=FALSE, warning=FALSE}
p + 
  scale_fill_GEMINI() +
  theme_GEMINI()
```
## Univariate Plot Function

The `eda_univ_autoplot` function creates univariate plots based on the variable type. The standard behavior is that the function will try to guess what is the data type of the variable supplied. That can be controlled by the user with the argument `input_type`, where the user can choose between `"numeric"` and `"character"`.

### Numeric variable
```{r, message=FALSE, warning=FALSE}
eda_univ_autoplot(mtcars, "mpg")
```
### Categoric variable
```{r, message=FALSE, warning=FALSE}
mtcars$cyl <- factor(mtcars$cyl)
eda_univ_autoplot(mtcars, "cyl")
```

### User-defined variable type
```{r, message=FALSE, warning=FALSE}
eda_univ_autoplot(mtcars, "mpg", input_type = "categorical")
```

## Bivariate Plot Function

The `eda_biv_autoplot` function creates bivariate plots based on the types of the covariate and outcome variables. Similarly to the `eda_univ_autoplot`, the user can define the type of the variables using the `input_type_out`, `input_type_cov` arguments. If not specified, the function will guess what the type of the data is.
### Numeric x Numeric
```{r, message=FALSE, warning=FALSE}
eda_biv_autoplot(mtcars, "wt", "mpg")
```
### Categorical vs. Categorical
```{r, message=FALSE, warning=FALSE}
mtcars$am <- factor(mtcars$am)
mtcars$cyl <- factor(mtcars$cyl)
eda_biv_autoplot(mtcars, "cyl", "am")
```
### Categorical vs. Numeric
```{r, message=FALSE, warning=FALSE}
eda_biv_autoplot(mtcars, "cyl", "mpg")
```
## Summary Table Functions

The package includes functions to generate summary tables for univariate and bivariate analyses.
### Univariate tabke
```{r, message=FALSE, warning=FALSE}
eda_univ_summary(mtcars, "mpg")
```

### Bivariate table
```{r, message=FALSE, warning=FALSE}
eda_biv_summary(mtcars, "cyl", "mpg")
```
# Conclusion

This vignette provided an overview of the `geminiEDA` package and demonstrated how to use its functions to create plots and generate summary tables for EDA. For more information, please refer to the package documentation.
